---
title: "Problem Set #4"
author: "Debi Gonzalez"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
# Using this library to read in the file.
nc_election <- read_csv("ps_4_elections-poll-nc09-3.csv")

# The following libraries are necessary to access the functions necessary such as spread, filter and summarize.
library(tidyverse)
library(devtools)
library(lubridate)
library(gt)
# The gt library is necessary for the table we are going to recreate in this PSET.
library(ggplot2)
# I am using this library to create my own plot(s). Yes, I made an extra one!


```


## Question 1

```{r, echo = FALSE}
# Number of respondent who supported democratic candidate
Dems <- nc_election %>%
  group_by(response) %>%
  # Response seems to provide what people answered when asked about their political support, so I will use that one.
  count() %>%
  spread(response, n) %>%
  pull(Dem)
  # There is probably a faster way to do this, but I counted to get the number of answers per response type, then I spread it so that the types of responses (Dem, Rep, Und, etc) become columns, and then I pull Dem to get the number of respondents who support democrats in the dataset.

# Difference between Republicans & Undecided
Reps <- nc_election %>%
  filter(response %in% c("Rep", "Dem", "Und")) %>%
  count(response) %>%
  spread(response, n) %>%
  summarize(Reps = Rep - Und)
  # I could also mutate to create the new variable and then pull the result. 
  # For this one, I added the filter just to make it more visually appealing to me and to get rid of irrelevant third party data, however, it is not necessary. Just like last time, I counted and spread to make the affiliations columns. Then, I summarized to create a new variable made by substracting the number of undecided respondents from the number of Republicans. This new value is the number of more republicans than undecided respondents.

# Number of individuals with different gender values
Gender <- nc_election %>%
  filter(gender != gender_combined)
  # This filter (!=) helps us to identify the individuals whose two gender variables differ in value. Using nrows() in inline code to refer to the number of individuals with different genders between the two variables by counting the number of rows in the tibble provided by this code. 

# Number of "White" respondents with race variables that differ in value
Race <- nc_election %>%
  filter(race_eth == "White") %>%
  filter(race_eth != file_race_black)
  # We filter for those respondents who are labeled "White" in the race_eth variable to reduce the amount of unnecessary data. Then, we filter for differences between race_eth and file_race_black (the other race variable) in order to identify the number of "White" respondents who have a different value for the file_race_black variable.

# Minute difference between the first Democrat and Republican respondents
Time <-  nc_election %>%
  filter(response %in% c("Rep", "Dem")) %>%
  group_by(response) %>%
  summarize(min_time = min(timestamp)) %>%
  spread(response, min_time) %>%
  mutate(time_diff = round(Rep - Dem, digits = 0)) %>%
  select(time_diff) 
  # Since I need to use the first timestamps of only republicans and democrats, I will filter for just them. Grouping by response helps organize the data and will allow me to see the first timestamp of Rep and Dem respectively when I summarize for the minimum timestamp (the smaller = first in terms of time). Spread fixes the table so that Rep and Dem become columns so that I can then create a new variable (time_diff) with the difference of the two columns with mutate. By trial and error I figured out that round works inside the mutate, but won't work with a variable outside the mutate itself. Digits = 0 because there are no digits after the decimal point, I want whole minutes. Finally, I select time_diff to isolate the difference in time response.
# Must find the min of both Reps and Dems 
  
```


There were `r Dems` respondents who supported the Democratic candidate.

There were `r Reps` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r nrow(Gender)` individuals for whom these variables have different values.

There are `r nrow(Race)` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Dem came `r Time` minutes (rounded to the nearest minute) before the first response of Rep. 

Note: Fixed the typo per Dillon's instruction. 


## Question 2
```{r, echo = FALSE, results="asis"}
nc_election %>%
  select(response, race_eth, final_weight) %>%
  # These are the variables necessary to replicate the table. Final weight is important because it will help us get the proper proportions of race to political identity/affiliation.
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  # We don't need those who did not have a classified race, so we will filter them out.
  mutate(race_eth = fct_relevel(race_eth, c("White", "Black", "Hispanic", "Asian", "Other"))) %>%
  # This mutate is to achieve the (race) list order of the model table.
  group_by(response, race_eth) %>%
  # Grouping in this order will help me spread the table later when I create the sum of final_weight variable.
  summarize(total = sum(final_weight)) %>%
  # Need to create this total variable in order to calculate the weighted number/proportions of response by race.
  spread(key =  response, value = total, fill = 0) %>% 
  # Spread so that our chart shows race in the first column and the response categories as the following set of columns. We use fill = 0 so that NA values can be calculated and weighted as well. 
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  # This chunk of code creates the proportions from the total. It creates a set of variables that we can use to calculate the exact weighted proportions of our data.
  select(-all, -`3`) %>% 
  # This diselects the all and third party columns from the table we are about to make so that it does not show.
  ungroup() %>% 
  # Will now explicitly ungroup so that I don't carry my grouping settings (which facilitated the making of proportions) into the final chart. 
  na_if(0) %>%
  # Using this function to get rid of the 0 in the proportion of Asian undecided voters.
  gt() %>% 
  # As we learned in class, this function will help us format out tibble into a pretty chart.
  fmt_percent(columns = vars(Dem, Rep, Und), decimals = 0) %>%
  # Using fmt_percent and specifying the variables we want to use percentage for converts our proportions to percentage.
  cols_label(race_eth = "", Dem = "DEM.", Rep = "REP.", Und = "UND.") %>%
  tab_header(title = "Polling Results from North Carolina by Race") %>%
  tab_source_note(source_note = "Source: New York Times Upshot/Siena College 2018 live polls")
  # This chunk of code labels the columns to mirror those in the model graph, provides a title (which was originally missing), and a source. 

```

## Question 3
```{r, echo = FALSE, results="asis"}
nc_election %>%
  select(educ, final_weight) %>% 
  filter(educ != "[DO NOT READ] Refused") %>% 
  # Selected the variables needed for the graph and cleaned out the education variable from unnecessary data (those who did not provide education level). Using mutate and fct_relevel to order educ (on its axis) in a chronological/logical manner to mirror the model.
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
  # Now, we are ready to make the graph using ggplot. Since educ is a categorical variable, I'm assigning it to the x-axis. Using geom_violin and geon_jitter to mimic the graph provided (alpha = .4 and width = .2 are the most similar I could get to the model graph.) 
  ggplot(aes(x = educ, y= final_weight)) +
  geom_violin() + geom_jitter(alpha = .4, width = .2) +
  coord_flip() + 
  # Had to flip coordinates to mirror the model graph and also because of aesthetics and x-axis category labels (long).
  xlab(NULL) +
  ylab("Weight Given to Respondent in Calculating Poll Results") +
  labs(title = "More Educated Matter Less in North Carolina 9th",
       subtitle = "Poll gives more weight to people who are less likely to participate in polls",
       caption = "New York Times Upshot/Siena College 2018 live polls")
  # Using the above functions I was able to mirror the labels of the model graph.
```

## Question 4

```{r, echo = FALSE, results = TRUE}
# Distribution of Democrats and Republicans by Education Level by Gender.
nc_election %>%
  select(response, gender, educ, file_race) %>%
  # These are the variables I am going to use in my plot. I selected file_race as well because I was testing this same data visual set up but with race instead of gender.
  filter(educ != "[DO NOT READ] Refused", response %in% c("Rep", "Dem")) %>%
  # I'm trying to clean my data of respondents who did not respond to the education variable and since I'm comparing Rep and Dem, I will filter for just the two identifications.
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
  # This line will order my education level in a sort of chronological order, fascilitates the graph. 
  group_by(response, educ) %>%
  # Grouping by both response and education level will help more clealy see the patterns by respective category when graphed.
  ggplot(aes(x=educ, y = response, color = educ)) + 
  geom_jitter(show.legend = FALSE) + 
  coord_flip() + 
  facet_wrap(~gender) +
  xlab(NULL) +
  ylab("Political Affiliation") +
  labs(title = "Political Affiliation by Education and Gender",
       caption = "Source: New York Times Upshot/Siena College 2018 live polls")
  # Chose jitter because out of all the plots I tried it best showed the distribution of my data by categories. Since I decided to facet_wrap by gender (only two side by side graphs), I colored by education to aid in the visualization of the distributions. The coordinate flip provided the most aesthetically pleasant version of the graph given that the education categories are long and would overlap at the bottom and the jitter plot would not show (as) significant data. I hid the legend because the graph is pretty self explanatory. I labeled the y axis (remember it's inverted), gave it a title and provided the source.

# I decided the gender variable was more interesting because I can see how political affiliation vary by gender with increasing education level. Females with higher education more strongly identified as Democrats. While males with graduate degrees seem only slightly more likely to identify as Republican, those with some college or trade school are more likely to be republican than their female countrparts with the same education level.
```

## Additional Plots (just for fun)
```{r, echo = FALSE, results= TRUE}
# Distribution of Average Timestamp by Gender (jitter)
nc_election %>%
  select(gender, timestamp) %>%
  group_by(gender, timestamp) %>%
  ggplot(aes(x=gender, y = timestamp, color =gender)) + geom_jitter(show.legend = FALSE) +
  xlab("Gender") +
  ylab("Day of Response Submission") +
  labs(title = "Timing of Response Submissions by Gender",
       caption = "New York Times Upshot/Siena College 2018 live polls")
```

```{r, echo = FALSE, results="asis"}
# Distribution of Average Timestamp by Gender (boxplot)
nc_election %>%
  select(gender, timestamp) %>%
  group_by(gender, timestamp) %>%
  ggplot(aes(x=gender, y = timestamp, color =gender)) + geom_boxplot(show.legend = FALSE) +
  xlab("Gender") +
  ylab("Day of Response Submission") +
  labs(title = "Timing of Response Submissions by Gender",
       caption = "New York Times Upshot/Siena College 2018 live polls")
```


## Colleagues

Diego M., Chris O., Chris M., Beau M., & Dillon S.