---
title: "Problem Set #4"
author: "Debi Gonzalez"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)

nc_election <- read_csv("ps_4_elections-poll-nc09-3.csv")



```


## Question 1

```{r}
Dems <- nc_election %>%
  filter(response %in% c("Rep", "Dem", "Und")) %>%
  group_by(response) %>%
  count() %>%
  spread(response, n) %>%
  pull(Dem)


Reps <- nc_election %>%
  filter(response %in% c("Rep", "Dem", "Und")) %>%
  count(response) %>%
  spread(response, n) %>%
  summarize(Reps = Rep - Und) %>%
  pull(Reps)
  # I could also mutate to create the new variable and then pull the result. 

Gender <- nc_election %>%
  filter(response %in% c("Rep", "Dem", "Und")) %>%
  filter(gender != gender_combined)
  # Will use nrows() in inline code to refer to the number of individuals with different genders between the two variables. 


Race <- nc_election %>%
  filter(response %in% c("Rep", "Dem", "Und")) %>%
  filter(race_eth == "White") %>%
  filter(race_eth != file_race_black) %>%
  nrow()
  

Time <-  nc_election %>%
  filter(response %in% c("Rep", "Dem")) %>%
  group_by(response) %>%
  summarize(min_time = min(timestamp)) %>%
  spread(response, min_time) %>%
  summarize(time_diff = Rep - Dem) %>%
  pull(time_diff) %>%
  round(time_diff, digits = 2)
  # round(time_diff, digits = 2)
 
  # 11

# Must find the min of both Reps and Dems 
  
  
```


There were `r Dems` respondents who supported the Democratic candidate.

There were `r Reps` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r nrow(Gender)` individuals for whom these variables have different values.

There are `r Race` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Rep came X minutes (rounded to the nearest minute) before the first response of Dem.